---
title: "AAPL Stock Data Analysis"
output: html_notebook
---

### Step 1: Initialize Environment

This chunk verifies and loads the quantitative modeling framework library.

```{r setup, message=FALSE, warning=FALSE}
# Automatically install quantmod if missing from your system
if(!require(quantmod)) install.packages("quantmod")

library(quantmod)
```

### Step 2: Fetch Financial Market Data

We use Yahoo Finance as our source endpoint. The end constraint date is configured as `2024-01-01` to safely encompass all trading sessions throughout December 2023.

```{r fetch_data}
getSymbols('AAPL', 
           src = 'yahoo', 
           from = '2023-01-01', 
           to = '2024-01-01')
```

### Step 3: Inspect Data Rows

Preview the top rows of the newly created extensible time-series (`xts`) object matrix.

```{r preview}
head(AAPL)
```

### Step 4: Visual Charting (Optional)

Run this block to render a candlestick price tracking profile directly inside your notebook output layout window.

```{r plot}
chartSeries(AAPL, 
            theme = chartTheme('white'), 
            name = 'AAPL 2023 Performance Overview')
```
