---
title: "R Integration with Financial Data Sources"
subtitle: "Automating Live Financial Data Feeds"
output: html_notebook
---

# Introduction: What is R Integration?

R can directly connect to live external resources:

\- Stock Market Data

\- Economic Indicators

\- Financial APIs

\- Online Data Sources

**Why do analysts use this?** - Download data automatically with code instead of pointing-and-clicking. - Build completely reproducible research scripts. - Avoid slow, error-prone manual Excel downloads.

------------------------------------------------------------------------

# 1. quantmod: Stock Data & Traditional Charts

### Purpose

Used for downloading stock prices, creating native financial charts, and quickly applying traditional technical analysis indicators.

### Load Package

```{r message=FALSE, warning=FALSE}
library(quantmod)
```

### Download Apple Stock Data

```{r message=FALSE, warning=FALSE}
getSymbols("AAPL")
```

### View Data Structure

```{r}
head(AAPL)
```

### Plot Stock Prices Inline

```{r}
chartSeries(AAPL)
```

> **Key Point:** `quantmod` = Quick Financial Data + Native Financial Charts

------------------------------------------------------------------------

-----------------------------------

# 2. tidyquant: The Tidyverse for Finance

### Purpose

Combines the power of **Tidyverse** data manipulation with **Financial Data structures**. It turns financial data into data frames (`tibbles`) that work flawlessly with familiar `dplyr` and `ggplot2` syntax.

```         
```

```{r}
install.packages("tidyquant")
```

### Load Package

```{r message=FALSE, warning=FALSE}
library(tidyquant)
```

### Download Stock Data into a Tidy Data Frame

```{r message=FALSE, warning=FALSE}
apple <- tq_get(
  "AAPL",
  from = "2025-01-01"
)

head(apple)
```

### Calculate Average Closing Price using dplyr Syntax

```{r}
library(dplyr)
```

```{r}
apple %>%
  summarise(
    Avg_Close = mean(close)
  )
```

> **Key Point:** `tidyquant` = Comprehensive Financial Analysis using Tidyverse Workflows

------------------------------------------------------------------------

--------------------

# 3. httr: Connecting to Custom Financial Web APIs

### Purpose

Connects R to external Web APIs. This is essential for sourcing unstructured data not covered by financial packages, such as live exchange rates, cryptocurrency tickers, and macroeconomic databases.

### Load Package

```{r message=FALSE, warning=FALSE}
library(httr)
```

### Make a Live API Request (Free Exchange Rate API)

```{r}
response <- GET(
  "https://api.exchangerate-api.com/v4/latest/USD"
)

response
```

### Convert Raw API JSON Response to Text

```{r}
content(response, "text")
```

> **Key Point:** `httr` = Connect R directly to any custom web API or data stream

------------------------------------------------------------------------

```{r}
library(httr)
library(jsonlite)

response <- GET(
  "https://api.exchangerate-api.com/v4/latest/USD"
)

data <- fromJSON(content(response, "text"))

str(data)
```

# Core Concept Workflow

### The Traditional Old Way (Inefficient)

``` text
Website → Manually Download CSV → Open Excel → Copy/Paste/Clean → Import into R
```

### The Automated R Way (Productive)

``` text
R Script → Automatically Fetch Data → Analyze Instantly → Live Visualization
```

------------------------------------------------------------------------

# Lesson Summary

| Package | Core Technical Purpose |
|:-----------------------------------|:-----------------------------------|
| **`quantmod`** | Quick market data scraping and immediate native charting. |
| **`tidyquant`** | Converts market data into neat data frames for Tidyverse flows. |
| **`httr`** | Connects R directly to custom HTTP REST APIs (Crypto, FX, Economics). |

**Why should a quantitative analyst use tools like quantmod, tidyquant, and httr?** \> *Answer:* These packages allow R to bypass manual downloads and pipe web-hosted financial data directly into memory. This eliminates Excel manual errors, automates recurring data refreshes, and ensures the entire analytics workflow is perfectly reproducible.
