---
title: "Financial Analytics in R"
---

# 1. Tidyverse: Data Wrangling & Transformation

### What is Tidyverse?

Tidyverse is a collection of packages used for: - Reading data - Cleaning data - Filtering records - Creating new variables - Summarizing data

Common packages: - `readr` $\rightarrow$ Import data - `dplyr` $\rightarrow$ Manipulate data - `tidyr` $\rightarrow$ Reshape data

### Load Package

```{r}
library(tidyverse)
```

### Create Sample Financial Dataset

```{r}
stocks <- data.frame(
  Company = c("TCS","Infosys","Wipro","HCL","TechM"),
  Price = c(3500, 1600, 450, 1200, 1400),
  Volume = c(10000, 15000, 8000, 12000, 9000)
)

stocks
```

### Select Columns

```{r}
stocks %>%
  select(Company, Price)
```

### Filter Expensive Stocks

```{r}
stocks %>%
  filter(Price > 1500)
```

### Create New Column

Suppose market value = Price $\times$ Volume

```{r}
stocks %>%
  mutate(MarketValue = Price * Volume)
```

### Sort by Price

```{r}
stocks %>%
  arrange(desc(Price))
```

### Summary Statistics

```{r}
stocks %>%
  summarise(
    AveragePrice = mean(Price),
    TotalVolume = sum(Volume)
  )
```

### Key Point: dplyr makes code readable

```{r}
stocks %>%
  filter(Price > 1000) %>%
  arrange(desc(Price))
```

*Read as: Take stocks* $\rightarrow$ Filter $\rightarrow$ Sort

------------------------------------------------------------------------

------------------------------

# 2. Quantmod: Quantitative Financial Modelling

### What is Quantmod?

Quantmod helps us: - Download stock prices - Create stock charts - Calculate indicators - Analyze trends

### Install (Once)

```{r eval=FALSE}
install.packages("quantmod")
```

### Load Package

```{r}
library(quantmod)
```

### Download Apple Stock Data

```{r message=FALSE, warning=FALSE}
getSymbols("AAPL")
```

*This downloads Apple stock prices.*

### View Data

```{r}
head(AAPL)
```

### Basic Stock Chart

```{r}
chartSeries(AAPL)
```

### Add Moving Average

Moving Average helps identify trends.

```{r}
chartSeries(AAPL)
addSMA(20, col="blue")
```

*20-Day Moving Average appears on chart.*

### Calculate Returns

```{r}
daily_returns <- dailyReturn(Cl(AAPL))
head(daily_returns)
```

*Where `Cl(AAPL)` means Closing Price.*

### Finance Interpretation

If return = `0.02`, then the stock gained **2%** for that day.

------------------------------------------------------------------------

------------------------

# 3. ggplot2: Professional Visualization (5 Minutes)

### What is ggplot2?

Used for: - Financial dashboards - Trend analysis - Statistical graphics - Business reporting

### Load Package

```{r}
library(ggplot2)
```

### Sample Stock Prices

```{r}
stock_data <- data.frame(
  Day = 1:10,
  Price = c(100,105,110,108,115,120,118,125,130,135)
)

stock_data
```

### Line Chart

```{r}
ggplot(stock_data, aes(x = Day, y = Price)) +
  geom_line()
```

### Better Looking Chart

```{r}
ggplot(stock_data, aes(x = Day, y = Price)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  labs(
    title = "Stock Price Trend",
    x = "Trading Day",
    y = "Price"
  )
```

### Bar Chart (Trading Volume Example)

```{r}
volume_data <- data.frame(
  Company = c("TCS","Infosys","Wipro","HCL"),
  Volume = c(10000,15000,8000,12000)
)

ggplot(volume_data, aes(x = Company, y = Volume)) +
  geom_bar(stat = "identity")
```

### Scatter Plot (Price vs Volume)

```{r}
ggplot(stocks, aes(x = Volume, y = Price)) +
  geom_point(size = 4)
```

------------------------------------------------------------------------

# Final Summary Slide

### Tidyverse

✔ Clean and transform financial data: `filter()`, `select()`, `mutate()`, `summarise()`

### Quantmod

✔ Download stock prices: `getSymbols()`

✔ Technical indicators: `addSMA()`, `dailyReturn()`

✔ Financial charts: `chartSeries()`

### ggplot2

✔ Professional visualizations: `geom_line()`, `geom_bar()`, `geom_point()`

✔ Financial dashboards and reports

### 
