---
title: "Financial Forecasting in R"
subtitle: "Time-Series Prediction with the forecast Package"
output: html_notebook
---

# 1. Forecast: Time-Series Prediction

### What is Forecasting?

Forecasting uses historical data to predict future values.

In Finance, forecasting is used for: - Stock Prices - Interest Rates - Exchange Rates - Sales Revenue - Economic Indicators

Popular Forecasting Models: - ARIMA - ETS (Exponential Smoothing) - Moving Average

### Install Required Package (Once)

```{r eval=FALSE}
install.packages("forecast")
```

### Load Required Package

```{r message=FALSE, warning=FALSE}
library(forecast)
```

### Create Sample Financial Data

Suppose this represents a stock's monthly closing price.

```{r}
stock_prices <- c(
  100,102,105,107,110,112,
  115,117,120,123,125,128
)

stock_prices
```

### Convert to Time Series

```{r}
stock_ts <- ts(stock_prices,
               start = c(2025,1),
               frequency = 12)

stock_ts
```

### Visualize the Trend

```{r}
plot(stock_ts,
     main="Monthly Stock Prices",
     ylab="Price",
     xlab="Month")
```

**Discussion Point for Students:** \> *Question:* Is the stock moving upward or downward? \> *Answer:* Upward trend.

------------------------------------------------------------------------

-------------------------

# 2. ARIMA Forecasting

### What is ARIMA?

ARIMA learns patterns from historical data and predicts future values. It is widely used in stock analysis, interest rate forecasting, and economic modeling.

### Build ARIMA Model

R automatically selects the best ARIMA parameters when using `auto.arima()`.

```{r}
arima_model <- auto.arima(stock_ts)

arima_model
```

### Predict Next 6 Months

```{r}
arima_forecast <- forecast(arima_model, h = 6)

arima_forecast
```

### Plot Forecast

```{r}
plot(arima_forecast, main="ARIMA Stock Price Forecast")
```

**Interpretation Key:**

\- **Black Line:** Historical prices

 - **Blue Line Extension:** Predicted future stock prices

\- **Shaded Region:** Uncertainty/confidence interval

------------------------------------------------------------------------

-----------------------

# 3. ETS Forecasting

### What is ETS?

ETS stands for **Error, Trend, and Seasonality**. It is another popular exponential smoothing framework for forecasting.

### Build ETS Model

```{r}
ets_model <- ets(stock_ts)

ets_model
```

### Forecast Next 6 Months

```{r}
ets_forecast <- forecast(ets_model, h = 6)

ets_forecast
```

### Plot ETS Forecast

```{r}
plot(ets_forecast, main="ETS Stock Price Forecast")
```

------------------------------------------------------------------------

------------------

# 4. Compare ARIMA and ETS

```{r}
plot(arima_forecast, main="ARIMA vs ETS")
lines(ets_forecast$mean, col="red", lwd=2)
```

**Interpretation Key:**

\- **Blue Line:** ARIMA prediction

 - **Red Line:** ETS prediction

\- *Note:* Different models can produce slightly different mathematical trajectories based on how they weight historical data.

------------------------------------------------------------------------

# 5. Real Financial Example: Interest Rates

### Create Interest Rate Dataset

```{r}
interest_rates <- c(
  5.0,5.1,5.2,5.3,5.4,5.5,
  5.6,5.7,5.8,5.9,6.0,6.1
)

rate_ts <- ts(interest_rates, frequency = 12)
```

### Forecast Future Interest Rates

```{r}
rate_model <- auto.arima(rate_ts)
rate_forecast <- forecast(rate_model, h = 6)

plot(rate_forecast, main="Interest Rate Forecast")
```

------------------------------------------------------------------------

# 6. Forecast Accuracy

How well does the model fit past data? We use evaluation metrics to check performance.

```{r}
accuracy(arima_forecast)
```

**Common Metrics:** - **MAE:** Mean Absolute Error - **RMSE:** Root Mean Square Error - *Rule:* Smaller error values indicate better, tighter mathematical predictions.

**Quick Business Interpretation:** If your forecast output shows: - Month 13: 130 - Month 14: 132 - Month 15: 134

> *Takeaway:* Based purely on historical trends, the asset value is structurally expected to continue its upward expansion over the next quarter.

------------------------------------------------------------------------

# Final Lesson Summary

### ARIMA

✔ Learns patterns from historical data (`auto.arima()`, `forecast()`)

 - Used for volatile market assets, stock prices, and economic indicators.

### ETS

✔ Models error, trend, and seasonality directly (`ets()`, `forecast()`)

\- Excellent for smooth, seasonal business trends and revenue forecasting.

### Visualizing Forecasts

✔ `plot(forecast_object)` plots history, predictions, and confidence risk blocks instantly.
