# AirPassengers Dataset in R

### Equivalent to Seaborn's`flights` Dataset

The built-in `AirPassengers` dataset in R contains monthly airline passenger counts from 1949 to 1960.

------------------------------------------------------------------------

## 1. Load the Dataset

```{r}
data(AirPassengers)
```

------------------------------------------------------------------------

## 2. View the Dataset

```{r}
AirPassengers
```

------------------------------------------------------------------------

## 3. Basic Information

```{r}
class(AirPassengers)
length(AirPassengers)
start(AirPassengers)
end(AirPassengers)
frequency(AirPassengers)
```

------------------------------------------------------------------------

## 4. Plot the Time Series

```{r}
plot(
  AirPassengers,
  main = "Monthly Airline Passengers (1949–1960)",
  xlab = "Year",
  ylab = "Passengers"
)
```

------------------------------------------------------------------------

## 5. Summary Statistics

```{r}
summary(AirPassengers)
mean(AirPassengers)
sd(AirPassengers)
min(AirPassengers)
max(AirPassengers)
```

------------------------------------------------------------------------

## 6. Convert to a Data Frame (Like Seaborn Flights)

```{r}
flights <- data.frame(
  year = rep(1949:1960, each = 12),
  month = factor(month.abb, levels = month.abb),
  passengers = as.numeric(AirPassengers)
)

head(flights)
tail(flights)
```

------------------------------------------------------------------------

## 7. Check Structure

```{r}
str(flights)
summary(flights)
```

------------------------------------------------------------------------

## 8. Scatter Plot: Year vs Passengers

```{r}
plot(
  flights$year,
  flights$passengers,
  pch = 19,
  xlab = "Year",
  ylab = "Passengers",
  main = "Passengers Over Time"
)
```

------------------------------------------------------------------------

## 9. Color Points Automatically by Month

```{r}
plot(
  flights$year,
  flights$passengers,
  col = as.numeric(flights$month),
  pch = 19,
  xlab = "Year",
  ylab = "Passengers",
  main = "Passengers Colored by Month"
)

legend(
  "topleft",
  legend = levels(flights$month),
  col = 1:12,
  pch = 19,
  cex = 0.7
)
```

------------------------------------------------------------------------

## 10. Average Passengers by Month

```{r}
aggregate(
  passengers ~ month,
  data = flights,
  FUN = mean
)
```

------------------------------------------------------------------------

## 11. Average Passengers by Year

```{r}
aggregate(
  passengers ~ year,
  data = flights,
  FUN = mean
)
```

------------------------------------------------------------------------

## 12. Boxplot by Month

```{r}
boxplot(
  passengers ~ month,
  data = flights,
  main = "Passenger Distribution by Month",
  xlab = "Month",
  ylab = "Passengers"
)
```

------------------------------------------------------------------------

## 13. Histogram of Passengers

```{r}
hist(
  flights$passengers,
  breaks = 20,
  main = "Distribution of Passengers",
  xlab = "Passengers"
)
```

------------------------------------------------------------------------

## 14. Find the Month with Highest Passengers

```{r}
flights[which.max(flights$passengers), ]
```

------------------------------------------------------------------------

## 15. Filter Data for Year 1955

```{r}
flights[flights$year == 1955, ]
```

------------------------------------------------------------------------

## 16. Filter Summer Months

```{r}
flights[
  flights$month %in% c("Jun", "Jul", "Aug"),
]
```

------------------------------------------------------------------------

## 17. Monthly Growth

```{r}
flights$growth <- c(NA, diff(flights$passengers))
head(flights, 10)
```

------------------------------------------------------------------------

## 18. Percentage Growth

```{r}
flights$pct_growth <- c(
  NA,
  diff(flights$passengers) /
    head(flights$passengers, -1) * 100
)

head(flights, 10)
```

------------------------------------------------------------------------

## 19. Month with Largest Increase

```{r}
flights[which.max(flights$growth), ]
```

------------------------------------------------------------------------

## 20. Save to CSV

```{r}
write.csv(
  flights,
  "flights.csv",
  row.names = FALSE
)
```

------------------------------------------------------------------------

## 21. Mini Exercises

```{r}
# Average passengers
mean(flights$passengers)

# Maximum passengers
max(flights$passengers)

# Average passengers by month
aggregate(passengers ~ month, flights, mean)

# Plot with colors by month
plot(
  flights$year,
  flights$passengers,
  col = as.numeric(flights$month),
  pch = 19
)
```

------------------------------------------------------------------------

## Key Concepts Covered

-   Time series data
-   Converting `ts` objects to data frames
-   Summary statistics
-   Scatter plots
-   Automatic color mapping
-   Grouped summaries with `aggregate()`
-   Histograms and boxplots
-   Growth calculations
-   Filtering data

```         
```
