# Beginner’s Guide to Data Analysis in R with the `iris` Dataset

The `iris` dataset is one of the most famous datasets in data science and statistics. It contains measurements of 150 flowers from three species.

### Species Included

-   Setosa
-   Versicolor
-   Virginica

### Features Measured

-   Sepal Length
-   Sepal Width
-   Petal Length
-   Petal Width

This dataset is excellent for teaching:

-   Data exploration
-   Summary statistics
-   Filtering
-   Grouped analysis
-   Basic visualization

------------------------------------------------------------------------

# Step 1: Load the Dataset

```{r}
# Load the built-in dataset
data(iris)

# View first 6 rows
head(iris)
```

------------------------------------------------------------------------

# Step 2: Understand the Structure

```{r}
str(iris)
dim(iris)
colnames(iris)
```

### Output

-   150 rows
-   5 columns
-   4 numeric columns
-   1 categorical column (`Species`)

------------------------------------------------------------------------

# Step 3: View Basic Statistics

```{r}
summary(iris)
```

This gives:

-   Minimum
-   1st Quartile
-   Median
-   Mean
-   3rd Quartile
-   Maximum

------------------------------------------------------------------------

# Step 4: Inspect Individual Columns

```{r}
mean(iris$Sepal.Length)
sd(iris$Sepal.Length)
min(iris$Sepal.Length)
max(iris$Sepal.Length)
```

------------------------------------------------------------------------

# Step 5: Count Species

```{r}
table(iris$Species)
```

Each species has 50 observations.

------------------------------------------------------------------------

# Step 6: Access Rows and Columns

```{r}
# First row
iris[1, ]

# First 5 rows
iris[1:5, ]

# Species column
iris$Species

# First two columns
iris[, 1:2]
```

------------------------------------------------------------------------

# Step 7: Filter Data

```{r}
# All Setosa flowers
setosa <- iris[iris$Species == "setosa", ]

head(setosa)
```

------------------------------------------------------------------------

# Step 8: Multiple Conditions

```{r}
large_petals <- iris[
  iris$Petal.Length > 5 &
  iris$Species == "virginica",
]

head(large_petals)
```

------------------------------------------------------------------------

# Step 9: Sort Data

```{r}
# Sort by Petal Length
iris_sorted <- iris[order(iris$Petal.Length), ]

head(iris_sorted)
tail(iris_sorted)
```

------------------------------------------------------------------------

# Step 10: Add a New Column

```{r}
iris$Petal.Area <- iris$Petal.Length * iris$Petal.Width

head(iris)
```

------------------------------------------------------------------------

# Step 11: Grouped Means

```{r}
aggregate(
  cbind(Sepal.Length, Petal.Length) ~ Species,
  data = iris,
  mean
)
```

------------------------------------------------------------------------

# Step 12: Grouped Standard Deviations

```{r}
aggregate(
  Petal.Length ~ Species,
  data = iris,
  sd
)
```

------------------------------------------------------------------------

# Step 13: Correlation Matrix

```{r}
numeric_iris <- iris[, 1:4]

cor(numeric_iris)
```

------------------------------------------------------------------------

# Step 14: Scatter Plot

```{r}
plot(
  iris$Sepal.Length,
  iris$Petal.Length,
  main = "Sepal Length vs Petal Length",
  xlab = "Sepal Length",
  ylab = "Petal Length"
)
```

```{r}
species_colors <- as.numeric(iris$Species)

plot(
  iris$Sepal.Length,
  iris$Petal.Length,
  col = species_colors,
  pch = 19
)

```

------------------------------------------------------------------------

# Step 15: Boxplot by Species

```{r}
boxplot(
  Petal.Length ~ Species,
  data = iris,
  main = "Petal Length by Species"
)
```

------------------------------------------------------------------------

# Step 16: Histograms

```{r}
hist(
  iris$Sepal.Length,
  main = "Histogram of Sepal Length",
  xlab = "Sepal Length"
)
```

------------------------------------------------------------------------

# Step 17: Pairs Plot

```{r}
pairs(iris[, 1:4])
```

------------------------------------------------------------------------

# Step 18: Simple Rule-Based Classification

```{r}
iris$Prediction <- ifelse(
  iris$Petal.Length < 2,
  "setosa",
  "other"
)

table(Actual = iris$Species, Predicted = iris$Prediction)
```

------------------------------------------------------------------------

# Step 19: Save to CSV

```{r}
write.csv(iris, "iris_data.csv", row.names = FALSE)
```

------------------------------------------------------------------------

# Step 20: Mini Exercises

``` r
# 1. Average Petal Width
mean(iris$Petal.Width)

# 2. Number of Virginica flowers
sum(iris$Species == "virginica")

# 3. Maximum Sepal Length
max(iris$Sepal.Length)

# 4. Correlation between Petal Length and Petal Width
cor(iris$Petal.Length, iris$Petal.Width)
```

------------------------------------------------------------------------
