---
title: "Exploratory Data Analysis in R"
subtitle: "Bank Marketing Dataset — A Finance Student's Guide"
author: "Ashwani Kumar"
date: "`r Sys.Date()`"
output:
  html_notebook:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    highlight: tango
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,     # show R code
  message = FALSE,    # suppress package messages
  warning = FALSE,    # suppress warnings
  fig.width  = 9,
  fig.height = 5.5,
  fig.align  = "center"
)
```

------------------------------------------------------------------------

# Introduction

> **Dataset:** Bank Marketing (UCI / Kaggle)\
> **Goal:** Did a client subscribe to a term deposit? (`y` = yes / no)\
> **Context:** Phone-call marketing campaign data from a Portuguese bank.

This notebook walks through a complete Exploratory Data Analysis (EDA) workflow.\
By the end you will be able to:

-   Load and inspect a real dataset\
-   Understand variable types and summary statistics\
-   Clean and transform data\
-   Visualise distributions, relationships and patterns\
-   Ask and answer business questions with code

------------------------------------------------------------------------

# Setup — Loading Packages

R uses **packages** (libraries) to extend its base functionality.\
We will use the `tidyverse` (a collection of data-science packages) and a few extras.

```{r load-packages}
# Install packages only if they are not already installed
required_packages <- c("tidyverse", "skimr", "ggcorrplot", "scales", "knitr", "RColorBrewer")

for (pkg in required_packages) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    install.packages(pkg)
  }
}

# Load packages
library(tidyverse)    # ggplot2, dplyr, tidyr, readr, etc.
library(skimr)        # rich summary statistics
library(ggcorrplot)   # correlation heatmaps
library(scales)       # axis formatting (commas, percent)
library(knitr)        # nice tables
library(RColorBrewer) # colour palettes

# A clean ggplot theme for all plots
theme_set(theme_minimal(base_size = 13))
```

------------------------------------------------------------------------

# Loading the Data

The file uses **semicolons** as separators (common in European datasets).\
We tell R about this with `read_delim()`.

```{r load-data}
# ── IMPORTANT ──────────────────────────────────────────────────────────────
# Place bank.csv in your working directory, then run:
#   setwd("path/to/your/folder")   # or use Session > Set Working Directory
# ───────────────────────────────────────────────────────────────────────────

bank <- read_delim("data_bank.csv", delim = ";", show_col_types = FALSE)

# Peek at the first few rows
head(bank, 8)
```

```{r dimensions}
# How big is the dataset?
cat("Rows   :", nrow(bank), "\n")
cat("Columns:", ncol(bank), "\n")
```

------------------------------------------------------------------------

# Understanding the Variables

## Data Dictionary

| Variable  | Type        | Description                                       |
|----------------|----------------|-----------------------------------------|
| age       | numeric     | Client age (years)                                |
| job       | categorical | Type of job                                       |
| marital   | categorical | Marital status                                    |
| education | categorical | Education level                                   |
| default   | binary      | Has credit in default?                            |
| balance   | numeric     | Average yearly balance (euros)                    |
| housing   | binary      | Has housing loan?                                 |
| loan      | binary      | Has personal loan?                                |
| contact   | categorical | Contact communication type                        |
| day       | numeric     | Last contact day of month                         |
| month     | categorical | Last contact month                                |
| duration  | numeric     | Last contact duration (seconds)                   |
| campaign  | numeric     | Number of contacts in this campaign               |
| pdays     | numeric     | Days since last contact (-1 = never contacted)    |
| previous  | numeric     | Contacts before this campaign                     |
| poutcome  | categorical | Outcome of previous campaign                      |
| **y**     | **binary**  | **Target: subscribed a term deposit? (yes / no)** |

## Variable Types in R

```{r variable-types}
# glimpse() gives a compact overview of every column
glimpse(bank)
```

```{r convert-factors}
# Convert character columns to factors (R's category type)
# Factors are essential for modelling and ordered plots

categorical_vars <- c("job", "marital", "education", "default",
                      "housing", "loan", "contact", "month",
                      "poutcome", "y")

bank <- bank %>%
  mutate(across(all_of(categorical_vars), as.factor))

# Verify
glimpse(bank)
```

------------------------------------------------------------------------

# Summary Statistics

## Base R Summary

```{r base-summary}
summary(bank)
```

## Rich Summary with `skimr`

`skim()` gives a much more informative snapshot — histograms included!

```{r skim}
skim(bank)
```

------------------------------------------------------------------------

# Target Variable Analysis

Before diving into features, always understand your **outcome variable**.

```{r target-counts}
# Frequency table
bank %>%
  count(y) %>%
  mutate(pct = n / sum(n))
```

```{r target-bar, fig.height=4}
bank %>%
  count(y) %>%
  mutate(pct = n / sum(n),
         label = paste0(round(pct * 100, 1), "%")) %>%
  ggplot(aes(x = y, y = n, fill = y)) +
  geom_col(width = 0.5, show.legend = FALSE) +
  geom_text(aes(label = label), vjust = -0.4, size = 5, fontface = "bold") +
  scale_fill_manual(values = c("no" = "#e07070", "yes" = "#4caf7d")) +
  scale_y_continuous(labels = comma) +
  labs(
    title    = "Term Deposit Subscription Rate",
    subtitle = "Target variable distribution — class imbalance is evident",
    x        = "Subscribed?",
    y        = "Number of Clients"
  )
```

> **Business Insight:** Only \~11–12 % of clients subscribed. This **class imbalance** matters when building predictive models later.

------------------------------------------------------------------------

# Univariate Analysis — One Variable at a Time

## Numeric Variables

### Age Distribution

```{r age-hist}
ggplot(bank, aes(x = age)) +
  geom_histogram(binwidth = 5, fill = "#4a90d9", colour = "white") +
  geom_vline(xintercept = median(bank$age), linetype = "dashed",
             colour = "firebrick", linewidth = 0.9) +
  annotate("text", x = median(bank$age) + 2, y = Inf,
           label = paste("Median =", median(bank$age)),
           vjust = 1.5, hjust = 0, colour = "firebrick", size = 4.5) +
  labs(title = "Distribution of Client Age",
       x = "Age", y = "Count")
```

### Account Balance Distribution

```{r balance-hist}
# Balance has extreme outliers — log scale helps visualise
bank %>%
  filter(balance > 0) %>%          # log is undefined for 0 / negative
  ggplot(aes(x = balance)) +
  geom_histogram(bins = 60, fill = "#f5a623", colour = "white") +
  scale_x_log10(labels = comma) +
  labs(title = "Distribution of Account Balance (positive values, log scale)",
       x = "Balance (€, log scale)", y = "Count")
```

```{}
```

### Call Duration

```{r duration-box, fig.height=4}
ggplot(bank, aes(x = duration)) +
  geom_boxplot(fill = "#9b59b6", alpha = 0.7, outlier.colour = "grey50") +
  scale_x_continuous(labels = comma) +
  labs(title = "Boxplot: Last Call Duration (seconds)",
       x = "Duration (s)", y = NULL) +
  theme(axis.text.y = element_blank())
```

> **Key R concept — boxplot anatomy:**\
> Box = IQR (Q1 to Q3) \| Line = Median \| Whiskers = 1.5 × IQR \| Dots = Outliers

# Bivariate Analysis — Two Variables at a Time

## Numeric vs. Target

### Do subscribers have higher balances?

```{r balance-median-table}
bank %>%
  group_by(y) %>%
  summarise(
    n           = n(),
    Median_Bal  = median(balance),
    Mean_Bal    = round(mean(balance), 0),
    SD_Bal      = round(sd(balance), 0)
  ) %>%
  kable(caption = "Balance Statistics by Subscription Outcome")
```

### 
