# Step 1: Load the Built-in Dataset

```{r}
# Load the dataset
data(EuStockMarkets)

# View the first few rows
head(EuStockMarkets)
```

------------------------------------------------------------------------

# Step 2: Understand the Dataset Structure

```{r}
# Check the structure
str(EuStockMarkets)


```

```{r}
# Dimensions
dim(EuStockMarkets)

# Column names
colnames(EuStockMarkets)
```

### Expected Output

-   1860 observations (days)
-   4 columns (indices)

------------------------------------------------------------------------

# Step 3: Convert to a Data Frame

Working with data frames is easier for beginners.

```{r}
stocks <- as.data.frame(EuStockMarkets)

head(stocks)

```

------------------------------------------------------------------------

# Step 4: View Basic Information

```{r}
# First 10 rows
head(stocks, 10)

# Last 10 rows
tail(stocks, 10)

# Summary statistics
summary(stocks)

```

------------------------------------------------------------------------

# Step 5: What Does Each Column Mean?

```{r}
colnames(stocks)
```

Returns:

-   `DAX`
-   `SMI`
-   `CAC`
-   `FTSE`

Each value is the closing level of that stock index on a trading day.

------------------------------------------------------------------------

# Step 6: Access One Column

```{r}
# View DAX values
head(stocks$DAX)

# Mean DAX level
mean(stocks$DAX)

# Standard deviation
sd(stocks$DAX)
```

```{r}
# Minimum and maximum
min(stocks$DAX)
max(stocks$DAX)
```

------------------------------------------------------------------------

# Step 7: Calculate Summary Statistics for All Columns

```{r}
summary(stocks)

```

```{r}
sapply(stocks, mean)
```

```{r}
sapply(stocks, sd)
sapply(stocks, min)
sapply(stocks, max)
```

------------------------------------------------------------------------

# Step 8: Plot All Indices

```{r}
plot(EuStockMarkets)
```

This produces a line chart showing all four indices over time.

------------------------------------------------------------------------

# Step 9: Plot One Index

```{r}
plot(
  stocks$DAX,
  type = "l",
  main = "DAX Index",
  xlab = "Day",
  ylab = "Index Level"
)
```

------------------------------------------------------------------------

# Step 10: Compare Two Indices

```{r}
plot(
  stocks$DAX,
  type = "l",
  col = "blue",
  ylim = range(stocks$DAX, stocks$FTSE),
  main = "DAX vs FTSE",
  xlab = "Day",
  ylab = "Index Level"
)

lines(stocks$FTSE, col = "red")

legend(
  "topleft",
  legend = c("DAX", "FTSE"),
  col = c("blue", "red"),
  lty = 1
)
```

------------------------------------------------------------------------

# Step 11: Correlation Matrix

```{r}
cor(stocks)
```

This shows how strongly the markets move together.

------------------------------------------------------------------------

# Step 12: Scatter Plot

```{r}
plot(
  stocks$DAX,
  stocks$CAC,
  main = "DAX vs CAC",
  xlab = "DAX",
  ylab = "CAC"
)
```

------------------------------------------------------------------------

# Step 13: Create a Simple Portfolio

```{r}
# Equal-weighted portfolio
stocks$Portfolio <- rowMeans(stocks)

head(stocks)
```

------------------------------------------------------------------------

# Step 14: Plot the Portfolio

```{r}
plot(
  stocks$Portfolio,
  type = "l",
  main = "Equal Weighted Portfolio",
  xlab = "Day",
  ylab = "Portfolio Value"
)
```

------------------------------------------------------------------------

# Step 15: Identify Highest and Lowest Values

```{r}
which.max(stocks$DAX)
which.min(stocks$DAX)

max(stocks$DAX)
min(stocks$DAX)
```

------------------------------------------------------------------------

# Step 16: Daily Changes

```{r}
dax_changes <- diff(stocks$DAX)

head(dax_changes)

mean(dax_changes)
sd(dax_changes)
```

------------------------------------------------------------------------

# Step 17: Daily Returns (Simple)

```{r}
dax_returns <- diff(stocks$DAX) / head(stocks$DAX, -1)

head(dax_returns)
mean(dax_returns)
sd(dax_returns)
```

------------------------------------------------------------------------

# Step 18: Histogram of Returns

```{r}
hist(
  dax_returns,
  main = "Histogram of DAX Returns",
  xlab = "Daily Return"
)
```

------------------------------------------------------------------------

# Step 19: Find Days with Positive Returns

```{r}
sum(dax_returns > 0)
sum(dax_returns < 0)
```

------------------------------------------------------------------------

# Step 20: Save to CSV

```{r}
write.csv(stocks, "JUNK_eu_stocks.csv", row.names = FALSE)
```

------------------------------------------------------------------------

# Mini Exercise for Students

``` r
# 1. Find average FTSE level
mean(stocks$FTSE)

# 2. Find highest CAC value
max(stocks$CAC)

# 3. Compute correlation between DAX and CAC
cor(stocks$DAX, stocks$CAC)

# 4. Plot SMI
plot(stocks$SMI, type = "l")
```

------------------------------------------------------------------------

# Key Finance Concepts Introduced

| R Concept   | Finance Concept        |
|-------------|------------------------|
| Data frame  | Market dataset         |
| Mean        | Average index level    |
| SD          | Volatility             |
| Correlation | Diversification        |
| Returns     | Performance            |
| Histogram   | Return distribution    |
| Row means   | Equal-weight portfolio |

------------------------------------------------------------------------
