Click **Knit** or run chunks step-by-step while teaching

---

# 📄 Complete Teaching File (Copy Everything Below)

````r
---
title: "Operators in R - Teaching Notebook"
author: "Instructor"
output: html_document
---

# Introduction

Operators in R are symbols used to perform operations on variables and values.

We will learn:

- Arithmetic Operators
- Relational Operators
- Logical Operators
- Assignment Operators
- Calculator Mini Project

---

# Arithmetic Operators

Arithmetic operators perform mathematical calculations.

```{r}

a <- 10
b <- 3

a + b      # Addition
a - b      # Subtraction
a * b      # Multiplication
a / b      # Division
a ^ b      # Power
a %% b     # Modulus (remainder)
a %/% b    # Integer division

````

---

# Relational Operators

Relational operators compare values and return TRUE or FALSE.

```{r}

x <- 5
y <- 10


x == y
x != y
x > y
x < y
x >= 5
y <= 10

```

---

# Logical Operators

Logical operators combine multiple conditions.

```{r}
x <- 5
y <- 10

(x > 2) & (y > 5)   # AND
(x > 2) | (y < 5)   # OR
(x > 2)
!(x > 2)            # NOT

```

---

# Assignment Operators

Assignment operators assign values to variables.

```{r}
x <- 10
y = 5
20 -> z

x
y
z
```

---

# Combined Calculation Example

```{r}
a <- 15
b <- 4

result <- (a + b) * 2
result
```

---

# Practice Example 1

```{r}
num <- 20

num %% 3
num > 10
(num > 10) & (num < 30)
```

---

# Practice Example 2

Evaluate the following expression:

```{r}
(5 + 3) * 2 > 10
```

---

# Mini Project: Calculator Script

This script allows users to perform calculations interactively.

```{r}
num1 <- as.numeric(readline("Enter first number: "))
num2 <- as.numeric(readline("Enter second number: "))

cat("\nChoose Operation\n")
cat("1 Addition\n")
cat("2 Subtraction\n")
cat("3 Multiplication\n")
cat("4 Division\n")
cat("5 Power\n")
cat("6 Modulus\n")
cat("7 Integer Division\n")

choice <- as.numeric(readline("Enter choice (1-7): "))

if(choice == 1){
  cat("Result =", num1 + num2)
}

else if(choice == 2){
  cat("Result =", num1 - num2)
}

else if(choice == 3){
  cat("Result =", num1 * num2)
}

else if(choice == 4){
  if(num2 == 0){
    cat("Division by zero not allowed")
  } else {
    cat("Result =", num1 / num2)
  }
}

else if(choice == 5){
  cat("Result =", num1 ^ num2)
}

else if(choice == 6){
  cat("Result =", num1 %% num2)
}

else if(choice == 7){
  cat("Result =", num1 %/% num2)
}

else{
  cat("Invalid choice")
}
```

---

# Assignment for Students

1. Modify calculator to allow repeated calculations using loop.
2. Add square root operation.
3. Add input validation.

---

# Summary

In this lesson, we learned:

* Arithmetic Operators
* Relational Operators
* Logical Operators
* Assignment Operators
* Building a Calculator using R

```

---

# ✅ How You Can Use This in Class

### Step 1  
Save file as:
```

operators_in_r.Rmd

```

### Step 2  
Open in RStudio

### Step 3  
Run chunk-by-chunk while teaching

### Step 4  
Knit into HTML/PDF for students

---

If you want, I can also create:

✅ Quarto version (modern industry format)  
✅ PPT version of this content  
✅ Student practice worksheet  
✅ Live classroom exercises  
✅ GUI calculator using Shiny  
✅ Assessment + quiz  

Just tell 👍
```
